BART: A Comparison with Non-bayesian Tree Methods

### 2024-12-07 paper/paper.tex
commit with code on simulated data

 We run BART from \cite{mccullochBARTBayesianAdditive2024}; Boosting from \cite{ridgewayGbmGeneralizedBoosted2024}; Random Forest and Bagging from \cite{breimanRandomForestBreimanCutlers2024}. We use half of the sample to fit $\hat{y}$, that is the Age variable, and half the sample to give us out of bag Mean Squared Errors (MSE) for predicting the same variable. Models are run for various values of TT, and for each value, both the mean squared error (MSE) of out-of-sample predictions (shown in plot \ref{plot_mse}) and the computation time (depicted in plot \ref{plot_time}) are recorded.

@@

  \toprule
  Variable Name & Role & Type & Description & Units & Missing Values\\  
  \midrule
   Sex                  & Feature       & Categorical   & M, F, and I (infant) & -             & No                      \\   Length               & Feature       & Continuous    & Longest shell measurement & mm         & No                      \\   Diameter             & Feature       & Continuous    & Perpendicular to length & mm           & No                      \\   Height               & Feature       & Continuous    & With meat in shell      & mm           & No                      \\   Whole\_weight         & Feature       & Continuous    & Whole abalone           & grams        & No                      \\   Shucked\_weight       & Feature       & Continuous    & Weight of meat          & grams        & No                      \\   Viscera\_weight       & Feature       & Continuous    & Gut weight (after bleeding) & grams    & No                      \\   Shell\_weight         & Feature       & Continuous    & After being dried       & grams        & No                      \\   Rings                & Target        & Integer       & +1.5 gives the age in years & -          & No                      \\
   \bottomrule
  \end{tabular}
  \caption{Table of Variables}
  \label{table1}
  \end{table}

>>>>

@@

  \toprule
  Variable Name & Role & Type & Description & Units \\
  \midrule
  Sex                  & Feature       & Categorical   & M, F, and I (infant) & -    \\
  Length               & Feature       & Continuous    & Longest shell measurement & mm\\
  Diameter             & Feature       & Continuous    & Perpendicular to length & mm  \\
  Height               & Feature       & Continuous    & With meat in shell      & mm  \\
  Whole\_weight         & Feature       & Continuous    & Whole abalone           & gram\\
  Shucked\_weight       & Feature       & Continuous    & Weight of meat          & gram\\
  Viscera\_weight       & Feature       & Continuous    & Gut weight (after bleeding) & \\
  Shell\_weight         & Feature       & Continuous    & After being dried       & gram\\
  Rings                & Target        & Integer       & +1.5 gives the age in years & - \\

   \bottomrule
  \end{tabular}
  \caption{Table of Variables}
  \label{table1}
  \end{table}

We run BART from \cite{mccullochBARTBayesianAdditive2024}; Boosting from \cite{ridgewayGbmGeneralizedBoosted2024}; Random Forest and Bagging from \cite{breimanRandomForestBreimanCutlers2024}. We use half of the sample to fit $\hat{y}$, that is the Age variable, and half the sample to give us out of bag Mean Squared Errors (MSE) for predicting the same variable. Models are run for various values of T, and for each value, both the mean squared error (MSE) of out-of-sample predictions (shown in plot \ref{plot_mse}) and the computation time (depicted in plot \ref{plot_time}) are recorded.
